---
title: "Analyse Hackage Snapshot"
author: "Torsten Kemps-Benedix"
date: "9.2.2018"
output: html_document
params:
  run_date: !r as.Date("2018-02-14")
---

# Setup

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

dbGetQuery1 <- function(con, sql, ...) {
  data.table(dbGetQuery(con,
    sqlInterpolate(ANSI(), 
                   sql, 
                   run_date = paste0("'", as.character(params$run_date), "'"),
                   ...)))
}

psqlArray <- function(a) {
  x <- substr(a, start = 2, stop = 1000)
  x <- substr(x, start = 1, stop = nchar(x) - 1)
  strsplit(x, ",")
}
```

```{r libraries}
library(ggplot2)
library(DBI)
library(RPostgreSQL)
library(reshape)
library(data.table)
library(MASS)
```

# Load Data

```{r load and enhance data}
drv <- dbDriver("PostgreSQL")
tryCatch({
  con <- dbConnect(drv, host = "127.0.0.1", dbname = "postgres")
  p <- dbGetQuery1(con, 
                   "SELECT * FROM hackage.package_snapshot 
                    WHERE run_date = ?run_date;")
  d <- dbGetQuery1(con, 
                   "SELECT * FROM hackage.dependency 
                    WHERE run_date = ?run_date;")
  b <- dbGetQuery1(con, 
                   "SELECT * FROM hackage.build_status 
                    WHERE run_date = ?run_date;")
}, finally = {
  dbDisconnect(con)})

p[, `:=`(
  daysSinceUpload = as.integer(Sys.Date() - as.Date(p$uploaded)),
  categories1 = psqlArray(categories),
  categories1l = sapply(psqlArray(categories), function(s) {
    sort(gsub("\\\"", "", tolower(s)))}),
  numVersions1 = sapply(psqlArray(versions), length)
  )]

setkey(p, package_name, version, run_date)
setkey(d, package_name, version, run_date, package_name_dependency)
setkey(b, package_name, version, run_date, ghc_version)

p[, rating1 := ifelse(is.na(rating),0,rating)]
p[, sizeF := factor(ifelse(size < 1000, 1, ifelse(size > 1e5, 3, 2)))]

b1 <- b[, .(okForAnyGhcVersion = any(build_status == "OK"),
            okForLatestGhc = any(ifelse(ghc_version == "8.2", build_status == "OK", FALSE))),
        by = .(package_name, version, run_date)]
setkey(b1, package_name, version, run_date)
p1 <- b1[p[run_date >= as.Date("2018-01-30")]]
p1[, okForAnyGhcVersion := ifelse(is.na(okForAnyGhcVersion), FALSE, okForAnyGhcVersion)]
p1[, okForLatestGhc := ifelse(is.na(okForLatestGhc), FALSE, okForLatestGhc)]
```

We have found `r nrow(p)` packages in the snapshot of `r format(params$run_date)`. Most packages' current version was uploaded several years ago. The following diagram shows the time since upload for all packages. The blue line marks the median which is `r round(median(p1$daysSinceUpload/365.25), 2)` years and the green line shows the mean which is `r round(mean(p1$daysSinceUpload/365.25), 2)` years.

```{r}
ggplot(p1, aes(daysSinceUpload/365.25)) +
  geom_histogram(bins = 50) +
  xlab("years since upload") +
  geom_vline(xintercept = median(p1$daysSinceUpload/365.25), col = "blue") +
  geom_vline(xintercept = mean(p1$daysSinceUpload/365.25), col = "darkgreen")
```

```{r}
cats <- gsub("\\\"", "", sort(unique(unlist(p1$categories1))))
catsl <- unique(tolower(cats))
```

We currently have `r length(cats)` different categories. There are some cases where categories differ only by upper vs. lower case. We treat these as the same categories. After this clean up there remain `r length(catsl)` different categories.

```{r categories}
pc <- rbindlist(lapply(catsl,
                       function(c) p1[sapply(p1$categories1l,
                                             function(v) is.element(c, v)),
                                      .(package_name, version, run_date,
                                        downloads_total, downloads_30days,
                                        okForAnyGhcVersion, okForLatestGhc, 
                                        size, rating,
                                        category = c)]))
c <- pc[, .(no_packages = .N,
            category_downloads_total = sum(downloads_total),
            category_downloads_30days = sum(downloads_30days)),
        by = category][order(category_downloads_30days, decreasing = TRUE)]
pc <- merge(pc, c, by = "category")
kable(c[1:100])
```

Most important packages in each of the most important categories:

```{r}
nPackages <- 20
nCategories <- 40
pcc <- pc[order(category, downloads_30days, decreasing = TRUE)
          ][, .SD[1:nPackages], by = category
            ][order(category_downloads_30days, decreasing = TRUE),
              .(category, package_name, downloads_total, downloads_30days, 
                downloads_30daysPctCat = round(100*downloads_30days/category_downloads_30days,1),
                okForAnyGhcVersion, okForLatestGhc, sizeKB = round(size/1024,0), rating)]
kable(pcc[category %in% c[1:nCategories, category]])
```

